---
title: "Hands-on Exercise 1"
author: "Lau Jia Yi"
---

# 1 Geospatial Data Wrangling with R

## 1.1 Data Sources

-   Master Plan 2014 Subzone Boundary (Web) from [data.gov.sg](https://data.gov.sg/)

-   Pre-Schools Location from [data.gov.sg](https://data.gov.sg/)

-   Cycling Path from [LTADataMall](https://www.mytransport.sg/content/mytransport/home/dataMall.html)

-   Latest version of Singapore Airbnb listing data from [Inside Airbnb](http://insideairbnb.com/get-the-data.html)

## 1.2 R Packages used

-   **sf** for importing, managing, and processing geospatial data, and
-   **tidyverse** for performing data science tasks such as importing, wrangling and visualising data.
-   **readr** for importing csv data,
-   **readxl** for importing Excel worksheet,
-   **tidyr** for manipulating data,
-   **dplyr** for transforming data, and
-   **ggplot2** for visualising data

## 1.3 Loading of R Packages

```{r}
pacman::p_load(sf, tidyverse)
```

Note: p_load function of pacman package is used to install and load sf and tidyverse packages into R environment.

## 1.4 Importing Geospatial data

```{r}
mpsz = st_read(dsn = "data/geospatial", 
                  layer = "MP14_SUBZONE_WEB_PL")
```

# Plotting a simple bar chart

```{r}
ggplot(data=listing_data,
       aes(x=price)) +
  geom_bar()
```
