---
title: "In class Exercise 07"
author: "Lau Jia Yi"
date: October 14, 2024
date-modified: "last-modified"
execute:
  eval: true
  echo: true
  warning: false
  freeze: true
---

```{r}
pacman::p_load(olsrr, corrplot, ggpubr, sf, spdep, GWmodel, tmap, tidyverse, gtsummary, dplyr, sfdep)
```

```{r}
condo_resale <- read_csv("data/aspatial/Condo_resale_2015.csv")
```

```{r}
glimpse(condo_resale)
```

```{r}
condo_resale_sf <- st_as_sf(condo_resale,
                            coords = c("LONGITUDE", "LATITUDE"),
                            crs=4326) %>%
  st_transform(crs=3414)
```

```{r}
condo_mlr <- lm(formula = SELLING_PRICE ~ AREA_SQM +
                  AGE + PROX_CBD + PROX_CHILDCARE +
                  PROX_ELDERLYCARE + PROX_URA_GROWTH_AREA +
                  PROX_HAWKER_MARKET + PROX_KINDERGARTEN +
                  PROX_MRT + PROX_PARK + PROX_PRIMARY_SCH +
                  PROX_TOP_PRIMARY_SCH + PROX_SHOPPING_MALL +
                  PROX_SUPERMARKET + PROX_BUS_STOP +
                  NO_Of_UNITS + FAMILY_FRIENDLY +
                  FREEHOLD,
                data=condo_resale_sf)
```

### Model Assessment: olsrr method

```{r}
ols_regress(condo_mlr)
```

### Multi-collinearity

```{r}
ols_vif_tol(condo_mlr)
```

### Variable selection

```{r}
condo_fw_mlr <- ols_step_forward_p(
  condo_mlr,
  p_val = 0.05,
  details = FALSE)
```

### Test for Normality Assumption

Test the model directly via \$model call from condo fw mlr.

```{r}
ols_plot_resid_fit(condo_fw_mlr$model) 
```

```{r eval=FALSE}
mlr_output <- as.data.frame(condo_fw_mlr$model$residuals) %>%
  rename(`FW_MLR_RES` = `condo_fw_mlr$model$residuals`)
```

```{r eval=FALSE}
condo_resale_sf <- cbind(condo_resale_sf,
                         mlr_output$FW_MLR_RES) %>%
  rename(`MLR_RES` = `mlr_output.FW_MLR_RES`)
```

```{r eval=FALSE}
tmap_mode("view")
tm_shape(mpsz)+
  tmap_options(check.and.fix = TRUE) +
  tm_polygons(alpha = 0.4) +
tm_shape(condo_resale_sf) +
  tm_dots(col = "MLR_RES",
          alpha = 0.6,
          style = "quantile")
tmap_mode("plot")
```

```{r}
condo_resale_sf <- condo_resale_sf %>%
  mutate(nb = st_knn(geometry, k=6,
                    longlat = FALSE),
         wt = st_weights(nb,
                         style = "W"),
         .before = 1)
```

```{r eval=FALSE}
set.seed(1234)

global_moran_perm(condo_resale_sf$MLR_RES,
                  condo_resale_sf$nb,
                  condo_resale_sf$wt,
                  alternative = "two.sided",
                  nsim=99)
```

The Global Moran's I test for residual spatial autocorrelation shows that it's p-value is less than 0.00000000000000000000022 which is less than the alpha value of 0.05. Hence we will reject the null hypothesis that the residuals are randomly distributed.

Since the Observed Global moran I = 0.25586 which is greater than 0, we can infer that the residuals resemble cluster distribution.

```{r eval=FALSE}
bw_fixed <- bw.gwr(formula = SELLING_PRICE ~ AREA_SQM +
                         AGE + PROX_CBD + PROX_CHILDCARE +
                         PROX_ELDERLYCARE + PROX_URA_GROWTH_AREA +
                         PROX_HAWKER_MARKET + PROX_KINDERGARTEN +
                         PROX_MRT + PROX_PARK + PROX_PRIMARY_SCH +
                         PROX_TOP_PRIMARY_SCH + PROX_SHOPPING_MALL +
                         PROX_SUPERMARKET + PROX_BUS_STOP +
                         NO_Of_UNITS + FAMILY_FRIENDLY +
                         FREEHOLD,
                       data=condo_resale.sf,
                       bw=bw_fixed,
                       kernel = 'gaussian',
                       longlat = FALSE) 
```

```{r eval=FALSE}
gwr_fixed <- gwr.basic(formula = SELLING_PRICE ~ AREA_SQM +
                         AGE + PROX_CBD + PROX_CHILDCARE +
                         PROX_ELDERLYCARE + PROX_URA_GROWTH_AREA +
                         PROX_HAWKER_MARKET + PROX_KINDERGARTEN +
                         PROX_MRT + PROX_PARK + PROX_PRIMARY_SCH +
                         PROX_TOP_PRIMARY_SCH + PROX_SHOPPING_MALL +
                         PROX_SUPERMARKET + PROX_BUS_STOP +
                         NO_Of_UNITS + FAMILY_FRIENDLY +
                         FREEHOLD,
                       data=condo_resale.sf,
                       bw=bw_fixed,
                       kernel = 'gaussian',
                       longlat = FALSE)
```

### Take-home EX! Take note of this to tidy up dataset.x

```{r eval=FALSE}
gwr_adaptive_output <- as.data.frame(
  gwr_adaptive$SDF) %>%
  select(-c(2:15))
```

