---
title: "Hands on Exercise 5b"
author: "Lau Jia Yi"
date: September 23, 2024
date-modified: "last-modified"
execute:
  eval: true
  echo: true
  warning: false
---

# **10  Local Measures of Spatial Autocorrelation**

## **10.1 Overview**

Local Measures of Spatial Autocorrelation (LMSA) focus on the relationships between each observation and its surroundings, rather than providing a single summary of these relationships across the map. In this sense, they are not summary statistics but scores that allow us to learn more about the spatial structure in our data. The general intuition behind the metrics however is similar to that of global ones. Some of them are even mathematically connected, where the global version can be decomposed into a collection of local ones. One such example are Local Indicators of Spatial Association (LISA). Beside LISA, Getis-Ord’s Gi-statistics will be introduce as an alternative LMSA statistics that present complementary information or allow us to obtain similar insights for geographically referenced data.

In this hands-on exercise, you will learn how to compute Local Measures of Spatial Autocorrelation (LMSA) by using **spdep** package. By the end to this hands-on exercise, you will be able to:

-   import geospatial data using appropriate function(s) of **sf** package,

-   import csv file using appropriate function of **readr** package,

-   perform relational join using appropriate join function of **dplyr** package,

-   compute Local Indicator of Spatial Association (LISA) statistics for detecting clusters and outliers by using appropriate functions **spdep** package;

-   compute Getis-Ord’s Gi-statistics for detecting hot spot or/and cold spot area by using appropriate functions of **spdep** package; and

-   to visualise the analysis output by using **tmap** package.

## **10.2 Getting Started**

### **10.2.1 The analytical question**

In spatial policy, one of the main development objective of the local govenment and planners is to ensure equal distribution of development in the province. Our task in this study, hence, is to apply appropriate spatial statistical methods to discover if development are even distributed geographically. If the answer is **No**. Then, our next question will be “is there sign of spatial clustering?”. And, if the answer for this question is yes, then our next question will be “where are these clusters?”

In this case study, we are interested to examine the spatial pattern of a selected development indicator (i.e. GDP per capita) of Hunan Provice, People Republic of China.(https://en.wikipedia.org/wiki/Hunan)

### **10.2.2 The Study Area and Data**

Two data sets will be used in this hands-on exercise, they are:

-   Hunan province administrative boundary layer at county level. This is a geospatial data set in ESRI shapefile format.

-   Hunan_2012.csv: This csv file contains selected Hunan’s local development indicators in 2012.

### **10.2.3 Setting the Analytical Toolls**

Before we get started, we need to ensure that **spdep**, **sf**, **tmap** and **tidyverse** packages of R are currently installed in your R.

-   sf is use for importing and handling geospatial data in R,

-   tidyverse is mainly use for wrangling attribute data in R,

-   spdep will be used to compute spatial weights, global and local spatial autocorrelation statistics, and

-   tmap will be used to prepare cartographic quality chropleth map.

The code chunk below is used to perform the following tasks:

-   creating a package list containing the necessary R packages,

-   checking if the R packages in the package list have been installed in R,

    -   if they have yet to be installed, RStudio will installed the missing packages,

-   launching the packages into R environment.

```{r}
pacman::p_load(sf, spdep, tmap, tidyverse)
```

## **10.3 Getting the Data Into R Environment**

In this section, you will learn how to bring a geospatial data and its associated attribute table into R environment. The geospatial data is in ESRI shapefile format and the attribute table is in csv fomat.

### **10.3.1 Import shapefile into r environment**

The code chunk below uses [*st_read()*](https://r-spatial.github.io/sf/reference/st_read.html) of **sf** package to import Hunan shapefile into R. The imported shapefile will be **simple features** Object of **sf**.

```{r}
hunan <- st_read(dsn = "data/geospatial", 
                 layer = "Hunan")
```

### **10.3.2 Import csv file into r environment**

Next, we will import *Hunan_2012.csv* into R by using *read_csv()* of **readr** package. The output is R data frame class.

```{r}
hunan2012 <- read_csv("data/aspatial/Hunan_2012.csv")
```
